Predicting the Quality of CDRH3 Antibody Loop Structural Models

### 2022-10-06 paper.tex
Spell-checked

Apart from being the most structurally diverse, the CDR-H3 loop is
also the most important for antigen binding, being located at the
center of the binding site and forming the most contacts with the
antigen\cite{MacCallum1996}. It was demonstrated that
differences in this loop alone are sufficient to enable otherwise
identical antibodies to distinguish between various
antigens\cite{Xu2000}.

@@

As some features are in the form of amino acid names, these must be
encoded before they can be passed to a ML-model. The
encoding strategy often determines how efficiently the ML-model learns
and how much information can be extracted. Different strategies were
employed to represent \highlight{followed by}\lilian{???} BLOSUM62\cite{XXXX} and \highlight{NLF}\lilian{???}\cite{XXXX} encoding. The
\highlight{physiochemical encoding}\lilian{explain} strategy was implemented for all ML-models, being
the most effective. However, PCA-3 BLOSUM62\cite{XXXX}, a dimensionality-reduced
BLOSUM62 encoding method achieved comparable results.
Feature selection was conducted to improve the ML-model's learning
capacity. A high-dimensional feature dataset bears the risk of
introducing excessive noise, facilitating ML-model overfitting and can be
responsible for an overall decrease in ML-model performance and
stability. Each additional input feature forces the ML-model to
handle a more complex task, which consumes excess computational power
and time and provides more variables leading to overfitting of the ML-model.

@@

\subsection{Hyperparameter Optimization}
In the process of hyperparameter optimization, the configuration of
ML-model parameters which results in best performance is selected. This
is usually a computationally expensive and manual procedure.
In an effort to automate this process, a population was defined for
each ML-model type, so hyperparameter optimization could be conducted
automatically for each ML-model and seamlessly integrated into the full
ML-model creation process. Two different methods for hyperparameter
optimization were tested. The first was a hybrid approach of randomized
search and grid search; the second used a genetic algorithm for
optimization. The genetic algorithm was found to achieve slightly
better results and was employeded for optimizing all ML-models.

@@

\subsection{Dataset-splitting}
The final ML-model was evaluated using a test set, separated from the
training set at the start in a 30/70 split (lock-box principle)
\highlight{35}\lilian{What is this?}. The performance of all individual sub-ML-models of the first layer
was determined using stratified K-folds cross-validation (K=10) as the
dataset is imbalanced, being skewed towards lower RMSD values. The
method is different from normal K-folds cross validation as it uses
stratified sampling instead of random sampling.
\lilian{I think it is still random, it's just that it is selected to represent the imbalance properly}
This ensures each
class is represented, as the percentage of samples for each class is
preserved.
A validation set, usually used for testing during the optimization
stage was omitted in favour of stratified K-folds cross-validation
(k=10)\cite{Krstajic2014,Kohavi1995}.
\lilian{I don't really understand what you mean here. The stratified
  cross-validation is going to use a test set in each fold --- it's
  just that this is selected to fairly represent the inbalance in the
  training data. Normally the validation set is the set that is held
  back for final testing --- which you say you did.}

@@

\section{Discussion}
The results suggest that our classifier can differentiate
between well-modelled and less well-modelled CDR-H3 loop
structures. An MCC value of 0.99 was achieved, which underlines this
ability for accurate discrimination. Different methods for data
pre-processing, feature encoding, feature selection and hyperparameter
optimization were tested.
Feature encoding methods that were very high-dimensional
(one-hot-encoding, BLOSUM62, NLF) were found to be
unfavorable. Dimensionality reduction methods (Principal Component
Analysis (PCA), Independent Component Analysis (ICA), projection-based
methods e.g.\ t-SNE) were used on BLOSUM62 encoded matrices, which lead
to significant improvement. However, a physicochemical encoding
strategy was most effective.
The selection of features incorporated in the training set seemed to
be most important for effective learning. A multitude of methods were
tested. No one fit-for-all method for the different ML-models could be
found.\lilian{I'm not sure what this means.} However, for our top-layer classifier in our final ML-model
recursive feature elimination worked best.
A set of commonly used machine learning algorithms were tested, and
the best ML-models were incorporated into the final ensemble ML-model. A
stacked ML-model approach (consisting of 23 binary classifiers and a
single top-layer nominal classifier) was shown to outperform single
ML-models.
An MCC value of 0.99 was achieved for a classifier predicting whether
an input 3D-model has an RMSD value below 2\AA, 2\AA--4\AA\ or above
4\AA.

>>>>

Apart from being the most structurally diverse, the CDR-H3 loop is
also the most important for antigen binding, being located at the
centre of the binding site and forming the most contacts with the
antigen\cite{MacCallum1996}. It was demonstrated that
differences in this loop alone are sufficient to enable otherwise
identical antibodies to distinguish between various
antigens\cite{Xu2000}.

@@

As some features are in the form of amino acid names, these must be
encoded before they can be passed to a ML-model. The
encoding strategy often determines how efficiently the ML-model learns
and how much information can be extracted. Different strategies were
employed to represent \highlight{followed by}\lilian{???} BLOSUM62\cite{XXXX} and \highlight{NLF}\lilian{???}\cite{XXXX} encoding. The
\highlight{physiochemical encoding}\lilian{explain} strategy was implemented for all ML-models, being
the most effective. However, PCA-3 BLOSUM62\cite{XXXX}, a dimensionality-reduced
BLOSUM62 encoding method achieved comparable results.
Feature selection was conducted to improve the ML-model's learning
capacity. A high-dimensional feature dataset bears the risk of
introducing excessive noise, facilitating ML-model over-fitting and can be
responsible for an overall decrease in ML-model performance and
stability. Each additional input feature forces the ML-model to
handle a more complex task, which consumes excess computational power
and time and provides more variables leading to over-fitting of the ML-model.

@@

\subsection{Hyperparameter Optimization}
In the process of hyperparameter optimization, the configuration of
ML-model parameters which results in best performance is selected. This
is usually a computationally expensive and manual procedure.
In an effort to automate this process, a population was defined for
each ML-model type, so hyperparameter optimization could be conducted
automatically for each ML-model and seamlessly integrated into the full
ML-model creation process. Two different methods for hyperparameter
optimization were tested. The first was a hybrid approach of randomized
search and grid search; the second used a genetic algorithm for
optimization. The genetic algorithm was found to achieve slightly
better results and was employed for optimizing all ML-models.

@@

\subsection{Dataset-splitting}
The final ML-model was evaluated using a test set, separated from the
training set at the start in a 30/70 split (lock-box principle)
\highlight{35}\lilian{What is this?}. The performance of all individual sub-ML-models of the first layer
was determined using stratified K-folds cross-validation (K=10) as the
dataset is imbalanced, being skewed towards lower RMSD values. The
method is different from normal K-folds cross validation as it uses
stratified sampling instead of random sampling.
\lilian{I think it is still random, it's just that it is selected to represent the imbalance properly}
This ensures each
class is represented, as the percentage of samples for each class is
preserved.
A validation set, usually used for testing during the optimization
stage was omitted in favour of stratified K-folds cross-validation
(k=10)\cite{Krstajic2014,Kohavi1995}.
\lilian{I don't really understand what you mean here. The stratified
  cross-validation is going to use a test set in each fold --- it's
  just that this is selected to fairly represent the imbalance in the
  training data. Normally the validation set is the set that is held
  back for final testing --- which you say you did.}

@@

\section{Discussion}
The results suggest that our classifier can differentiate
between well-modelled and less well-modelled CDR-H3 loop
structures. An MCC value of 0.99 was achieved, which underlines this
ability for accurate discrimination. Different methods for data
pre-processing, feature encoding, feature selection and hyperparameter
optimization were tested.
Feature encoding methods that were very high-dimensional
(one-hot-encoding, BLOSUM62, NLF) were found to be
unfavourable. Dimensionality reduction methods (Principal Component
Analysis (PCA), Independent Component Analysis (ICA), projection-based
methods e.g.\ t-SNE) were used on BLOSUM62 encoded matrices, which lead
to significant improvement. However, a physicochemical encoding
strategy was most effective.
The selection of features incorporated in the training set seemed to
be most important for effective learning. A multitude of methods were
tested. No one fit-for-all method for the different ML-models could be
found.\lilian{I'm not sure what this means.} However, for our top-layer classifier in our final ML-model
recursive feature elimination worked best.
A set of commonly used machine learning algorithms were tested, and
the best ML-models were incorporated into the final ensemble ML-model. A
stacked ML-model approach (consisting of 23 binary classifiers and a
single top-layer nominal classifier) was shown to outperform single
ML-models.
An MCC value of 0.99 was achieved for a classifier predicting whether
an input 3D-model has an RMSD value below 2\AA, 2\AA--4\AA\ or above
4\AA.
